\section{SYSTEM CALLS}

\subsection*{DATA STRUCTURES}

\noindent B1: (10 marks)
Copy here the declaration of each new or changed `struct' or
`struct' member, global or static variable, `typedef', or
enumeration.  Identify the purpose of each in 25 words or less.

\noindent B2: (5 marks)
Describe how file descriptors are associated with open files.
Are file descriptors unique within the entire OS or just within a
single process?

There can be multiple file descriptors for the same open file as well as multiple file descriptors for the same open file within the same process. So the file descriptors are only unique within a single process as opposed to the whole OS. This is because each process has a list of file descriptors associated with it as opposed to a global collection of file descriptors.

\subsection*{ALGORITHMS}

\noindent B3: (5 marks)
Describe your code for reading and writing user data from the
kernel.

\noindent B4: (5 marks)
Suppose a system call causes a full page (4,096 bytes) of data
to be copied from user space into the kernel.  What is the least and the greatest possible number of inspections of the page table (e.g. calls to pagedir\_get\_page()) that might result?  What about for a system call that only copies 2 bytes of data?  Is there room for improvement in these numbers, and how much?

\noindent B5: (8 marks)
Briefly describe your implementation of the "wait" system call
and how it interacts with process termination.

In terms of the system call itself, it just calls process\_wait() as there is no need to map from pid to tid. In process\_wait(), first of all, the parent thread looks for the thread associated with the child tid. Then the thread pointer is checked if it does not exist (==NULL) or whether or not that the process actually is a child of the current process. This is done by looking for the tid in the list of child tids associated with the parent process. These are errors and so return -1.
If the process is a child on the parent process, it is then checked if wait() has been called on it before by the parent, it does this by checking the flag waited\_upon. Again, this is an error so -1 is returned. Otherwise, waited\_upon is set and then the parent process waits for if the child process has been initialised by waiting until the flag process\_init has been set. After this it busy waits until either the pagedir is equal to null or the thread is equal to null. If the thread is equal to the null then the thread has been terminated by the kernel and -1 is returned. Otherwise, the process has been exited normally and the exit status is returned.

\noindent B6: (5 marks)
Any access to user program memory at a user-specified address
can fail due to a bad pointer value.  Such accesses must cause the process to be terminated. System calls are fraught with such
accesses, e.g. a "write" system call requires reading the system
call number from the user stack, then each of the call's three
arguments, then an arbitrary amount of user memory, and any of
these can fail at any point.  This poses a design and
error-handling problem: how do you best avoid obscuring the primary
function of code in a morass of error-handling?  Furthermore, when
an error is detected, how do you ensure that all temporarily
allocated resources (locks, buffers, etc.) are freed?  In a
paragraph, describe the strategy or strategies you adopted for
managing these issues.

% TODO: Abstract error checking into separate helper functions depending on the
type of system call.

% TODO: Also add freeing temporarily allocated resources to source code


\subsection*{SYNCHRONIZATION}

\noindent B7: (5 marks)
The "exec" system call returns -1 if loading the new executable
fails, so it cannot return before the new executable has completed
loading.  How does your code ensure this?  How is the load
success/failure status passed back to the thread that calls "exec"?

% TODO: Describe use of a semaphore

In our code, we implemented a semaphore that is initialised at 0 and the parent thread calls down on the semaphore before passing the reponsibility to the executing thread. When the child thread has successfully or unsuccessfully loaded the file, it then calls up on that semaphore, so that the parent thread is unblocked only after the child thread has attempted to load the executable.

\noindent B8: (5 marks)
Consider parent process P with child process C.  How do you
ensure proper synchronization and avoid race conditions when P
calls wait(C) before C exits?  After C exits?  How do you ensure
that all resources are freed in each case?  How about when P
terminates without waiting, before C exits?  After C exits?  Are
there any special cases?

If wait is called before C exits, then the program busy waits by yielding until C is terminated either by the kernel or by an exit call.

% TODO: Busy waits if wait is called before C exits.


\subsection*{RATIONALE}

\noindent B9: (5 marks)
Why did you choose to implement access to user memory from the
kernel in the way that you did?

% TODO

\noindent B10: (5 marks)
What advantages or disadvantages can you see to your design
for file descriptors?

Since each process has its own set of file descriptors, our design is advantageous as each process will only have access to its own file descriptors as is specified in Task 2 specification. Look ups for files are also more efficient than if there was a global list of file descriptors for the OS as the process only has to look up its own file descriptors as opposed to all of them currently present. The way that the file descriptors are generated is by looking for gaps in the file descriptor list of the process and then inserting a new fd element into the list. This has the advantage of being fairly easy to program (increment until the fd does not match one in the list) and has a fairly low complexity of O(n). However, a problem with this approach is that the list of fds themselves need to be sorted before generating a new fd according to the algorithm.

\noindent B11: (2 marks)
The default tid\_t to pid\_t mapping is the identity mapping.
Did you change this? Why?

We decided not to change this mapping as we found that there was no need to as each process can only have one thread and the pid is determined by the tid of the thread associated with it. It also made it easier in certain functions for casting and in debugging as we always knew that the thread id would be the same as the process id.
